---
title: "Lab 5"
author: "Anna Jeffries"
date: "`r Sys.Date()`"
output: pdf_document 
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,message=FALSE, fig.align='center')
```

\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
library(tidyverse)
library(dplyr)
library(pander)
library(ggplot2)
library(GGally)
library(dslabs)
library(equatiomatic)
library(corrplot)
```

## 1. Figure out where the 1460-1379 = 81 rows of data are going when using `model.matrix`.

```{r tidy=TRUE, tidy.opts=list(width.cutoff=60)}
ameslist  <- read.table(
  'https://raw.githubusercontent.com/ajkirkpatrick/FS20/postS21_rev/classdata/ames.csv', 
                   header = TRUE,
                   sep = ',') 
GarageTemp <- model.matrix( ~ GarageType - 1, data=ameslist )

# sum(is.na(ameslist$GarageType))
ameslist_0 <- ameslist[!is.na(ameslist$GarageType),] # those pesky NAs

ameslist <- cbind(ameslist_0, GarageTemp) # yeet haw
ameslist$GarageOutside <- ifelse(ameslist$GarageTypeDetchd == 1 | ameslist$GarageTypeCarPort == 1, 1, 0)

ameslist <- ameslist[!is.na(ameslist$GarageOutside),] # yeeting those NAs
```

I am electing to drop these missing rows. Garage data aside, I doubt there's anything particularly special about the missing entries. It's not a trivial percentage of data we're losing, admittedly, but this is a low-stakes kind of context and I don't want to bother with a complicated imputation for a binary variable right now. 

## 2. Prune the data to 6-8 of the variables that are `type = int` about which you have some reasonable intuition for what they mean. 

Choose those that you believe are likely to be correlated with `SalePrice.` This must include the variable `SalePrice` and `GrLivArea.` Save this new dataset as `Ames`.  
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#glimpse(ameslist)

ames <- ameslist[c('SalePrice', 'GrLivArea', 'OverallQual', 'OverallCond','LotArea', 'YearBuilt','YearRemodAdd')]

```

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| **Variable**   | **Definition**              | **Value**     |
|--------|:-----------------|:--------|
| GrLivArea      | Total livable area | Positive integer |
| OverallQual      | Some score of house quality      |   Positive integer |
| OverallCond | Some score of house condition       |    Positive integer |
| LotArea      | Area of the lot in square feet | Positive integer |
| YearBuilt | Year the house was built      |    Positive integer |
| YearRemodAdd      | Year the remodel was added (if at all) | Positive integer |
| SalePrice      | Price of house sale      |   Positive integer |
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

## 3. Produce a scatterplot matrix.
\  
```{r tidy=TRUE, tidy.opts=list(width.cutoff=60)}
ggpairs(ames, 
        upper = list(continuous = wrap("cor", size = 2.5)),
        lower = list(continuous = wrap("points", alpha = 0.3,    size=0.5))) + 
        theme_minimal() + theme(axis.text = element_text(size = 5), 
                                strip.text.x = element_text(size = 5),
                                strip.text.y = element_text(size = 5, angle=0),
                                axis.text.x = element_text(angle=40, vjust=1, hjust=.8),
                                axis.text.y = element_text(vjust=1, hjust=.8)) 
        # what an absolute pain in the ass
```

## 4. Compute a matrix of correlations between these variables. 

Do the correlations match your prior beliefs? Briefly discuss the correlation between the chosen variables and SalePrice and any correlations between these variables.  
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
ggcorr(ames, method=c("everything",'pearson'), progress=FALSE, label=TRUE, label_size=3)
```
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
pander(cor(ames))
```

Overall, there are some pretty strong correlations among the variables. As expected, though, the strongest correlations are between `GRLivArea`, `OverallQual`, and `SalePrice`. There's also a fairly strong relationship between `YearBuilt` and `YearRemodAdd`. As noted previously, `YearRemodAdd` is a complete field. One could rhapsodise upon the general correlation between the year a house is built and the year a remodel is added (but this depends heavily on what the definition fo `YearRemodAdd` is), or perhaps theorise that some small number of realtors or developers owned and/or remodeled the homes in an en masse sort of fashion and used some kind of function to determine when a house needed to be remodeled. 

## 5. Produce a scatterplot between `SalePrice` and `GrLivArea`. Run a linear model using `lm()` to explore the relationship. Finally, use the `geom_abline()` function to plot the relationship that you’ve found in the simple linear regression.

What is the largest outlier that is above the regression line? Produce the other information about this house.  
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70), fig.height=4}
ggplot(ames, aes(x=GrLivArea, y=SalePrice, color=factor(OverallQual))) + geom_point(size=.5) + 
  geom_smooth(method=lm , color="red", se=T, size=.5) + 
  theme_minimal() + labs(color="OverallQual")
```
\  
```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
mod <- lm(SalePrice ~ GrLivArea, ames)
pander(summary(mod))
```

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
ggplot(ames, aes(x=GrLivArea, y=SalePrice, color=factor(OverallQual))) + geom_point(size=.5) + 
  theme_minimal() + labs(color="OverallQual") + 
  geom_abline(mod,slope=107.8, intercept=20042, color="red")
```
\  

Well, well, well, if `geom_abline()` doesn't give me the exact same results as `geom_smooth()`....  
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# ameslist[order(ameslist$SalePrice, ameslist$GrLivArea,decreasing=T),c("Id","GrLivArea","SalePrice")]

# pander(ameslist[ameslist$Id == 692,])
```
\  
Finding the "largest outlier" is a bit tricky here. Visually, we can observe the dots above the regression line in our graph that appear to boast the *largest selling prices* and we can index to find the relevant information from the dataset. *However*, in terms of actual "largest outlier," we'd want to look at the residuals to determine the "largest"....  
\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# which(mod$residuals == max(mod$residuals))
# pander(ameslist[c(850,899),])
```



## Bonus.

\  

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# table(ameslist$CentralAir )

ggplot(ameslist, aes(x=YearBuilt, y=SalePrice, color=factor(CentralAir))) + geom_point(size=.5) + 
  theme_minimal() + labs(color="Central Air?") + scale_color_manual(values=c('red', 'darkgreen'))
```

